---
title: "Reviews_labour_market_US"
author: "Jaroslaw Dremlik"
date: "15/01/2022"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Labour market short analysis and visualisation excersize

### 1. load packages
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
```

### 2. Including Plots

Download files and upload data frame
```{r}

# path <-'C:/Main_folder/Source_files/R_source_file/Labour_Market_US/'
# files <- c('alldata.csv',
#            'fulltimeAL.csv',
#            'fulltimeAT.csv',
#            'fulltimeBO.csv',
#            'fulltimeBOS.csv',
#            'fulltimeCHI.csv',
#            'fulltimeDC.csv',
#            'fulltimeLA.csv',
#            'fulltimeMA.csv',
#            'fulltimeMV.csv',
#            'fulltimeNY.csv',
#            'fulltimeRM.csv',
#            'fulltimeSD.csv',
#            'fulltimeSEA.csv',
#            'fulltimeSF.csv',
#            'fulltimeSU.csv'
#            )
# frame <- function(k){
#   return(read.csv(paste(path, files[k], sep = ''),
#                   na.strings = c('NA', 'NULL')))
# }
# 
# df_all <- frame(1)


frame <- read.csv('alldata.csv', sep=',')
df_all <- frame
```

### Basic data frame statistical summary
```{r}
summary(df_all)
```

### Check NAs
```{r}
is.na(df_all)
```

### Excluding NAs
```{r}
df_all <- na.omit(df_all)
```

### Checks NAs
```{r}
sum(is.na(df_all$reviews))
```

### Plotting general view of job reviews, plot no.: 1
```{r message=TRUE, warning=TRUE}
cutting_into_breaks = cut(df_all$reviews, breaks = seq(from = 0, to = 500, 
                                                       by = 100))
w = table(cutting_into_breaks)
names(w) <- c('0 to 50', '100 to 150', '200 to 250', '300 to 350','400 to 450')
plot(w, ylab = 'Number of rewiews', xlab = 'Breaks of frequency', col = 'blue', 
     lwd = 3)
title('General overview of number of job reviews')
```

### Extracting and adding new variable City do set of data frame named City
```{r}
city_1 <- str_split_fixed(df_all$location, ',', 2)
city_1
city_2 <- city_1[,-2]
city_2
df_all$city <- city_2
```

### Extracting and adding new variable State do set of data frame named State
```{r}
state_2 <- city_1[,-1]
state_2
state_3 <- gsub(' ', '', state_2)
state_3
state_4 <- strtrim(state_3, 2)
state_4
df_all$state <- state_4
```

### Plotting average of reviews by State in the US, plot no.: 2
```{r}
aver_state_reviews <- sort(tapply(df_all$reviews, df_all$state, mean), 
                           decreasing = TRUE)
barplot(aver_state_reviews, col='lightblue', xlab = 'States in the US')
title('Average reviews per state in the US')
```

### Plotting average of reviews by company TOP-5, plot no.: 3
```{r}

company <- sort(tapply(df_all$reviews, df_all$company, mean), decreasing = TRUE)
company_top_5 <- head(company, 5)
barplot(company_top_5, col = 'pink')
title('Average reviews number by company TOP 5')

```

### Plotting average of reviews by company BOTTON-5, plot no.: 4
```{r}
company <- sort(tapply(df_all$reviews, df_all$company, mean), decreasing = FALSE)
company_bottom_5 <- head(company, 5)
barplot(company_bottom_5, col = 'brown')
title('Average reviews number by company BOTTOM 5')
```

### plotting Walmart average review by City, plot no.: 5
```{r}

walmart <- df_all %>%
  select(city, company, reviews, state) %>%
  filter(company == 'Walmart')
  
walmart_cities <- walmart[,-2]

ggplot(data = walmart_cities, aes(x=city, y=reviews, colour = state)) + 
  geom_point(size = 5) + labs(title = 'Walmart average number of review by City')

```

### plotting Top 5 companies with highetrs number of reviews, plot no.: 6
```{r}
company_data_scientist <- df_all %>%
  select(company, position, reviews) %>%
  filter(position %in% 'Data Scientist') %>%
  group_by(reviews)

company_data_scientist
company_data_scientist_mean <- sort(tapply(company_data_scientist$reviews, 
                                           company_data_scientist$company, sum), 
     decreasing = TRUE)

top_5 <- head(company_data_scientist_mean, 5)
top_5
names(top_5) <- c('Amazon', 'SHMC', 'IBM', 'Verizon', 'ITL USA')
barplot(top_5, ylab = 'Reviews', xlab = 'Company', col = 'darkred')
title('Top five companies - reviews for position Data Scientist')

```


### plotting Number of reviews for position "Data Scientist" by US state, plot no.: 7
```{r}
data_scientist_1 <- df_all %>%
  select(state, position, reviews) %>%
  filter(position == 'Data Scientist') %>%
  group_by(state) %>%
  arrange(state, desc(reviews)) %>%
  summarise(reviews = sum(reviews)) 

ggplot(data_scientist_1, aes(x=state, y=reviews)) + 
  geom_col(col = 'black', fill = 'lightblue') +
  theme_replace() + 
  labs(x = 'states', y = 'reviews', 
       title = 'Number of reviews of Data Scientist positions by US state')

```

### ploting number of reviews main histogram all states in US, plot no.: 8
```{r}
facets_state <- df_all %>%
  select(reviews, state)

ggplot(data=facets_state, aes(reviews)) + geom_histogram(color = 'green', 
                                                         bins = 30)

```

### ploting number of reviews histograms breakdown by state in US, plot no.: 9
```{r}
facets_state <- df_all %>%
  select(reviews, state)

ggplot(data=facets_state, aes(reviews)) + geom_histogram(color = 'red', 
                                                         bins = 30) + facet_wrap(~state)
```


```{r}

```


